---
title: 'Day 3: Experimental design and hypothesis testing'
author: "Ricardo Ramirez [roramirezf@uni-heidelberg.de]"
date: "6/27/2022"
output: html_document
---

Following your work from yesterday, today we will have discussions regarding the experimental design of the experiments from the heart failure meta-analysis. 

1. Can you describe what's the original question so as to generate the data from the experimeent you chose?

2. Can you mention what are the distinct factor or levels inside your experiment?

3. What's the experimental unit from your data?

4. Why is it dangerous that researchers don't report full characteristics of the patient cohort they analyze?

5. Asssume you had to design your own data set, no limits in money, you can use multiple tissue banks, how would you approach this problem?

6. In which circumstances experimental design can be more or less controlled?

7. Pick a study that has sex information and perform a PCA, color the samples by heart failure condition and sex, what do you observe? How would you quantify this observation?

8. Merge two or more studies and perform a PCA on the expression matrix, what happens? 

9. Repeat the same task as above, but before performing a PCA, center and standardize each gene. You can use R's ```scale()``` function. What happens? Why do you think the results are so different?

10. How would you perform differential expression analysis in a single study? Can you define these question in terms of a hypothesis test? Any tests you know of? Pick 10 random genes and test if there's a difference in expression between heart failure patients and healthy donors.

11. Pick a single study and scan the differential expression analysis slot from the provided data set, bassed on our discussions, pick the top 10 up and downregulated genes and make a heatmap with them.

Use this instance to have useful packages

```{r, eval=F}
install.packages("tidyverse")
install.packages("BiocManager")
install.packages("R.utils")
install.packages("tidyverse")
install.packages("BiocManager")
install.packages("cowplot")
BiocManager::install("limma")
```

```{r, eval=F}
download.file("https://www.dropbox.com/s/o9q3c57cqyopuju/METAheart.rds?dl=1", 
              "METAheart.rds")
              
reheat <- readRDS("METAheart.rds")
```

